Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-09-05 chatbot_classification/aaai_2021/paper.tex
modified:   paper.tex   modified:   picture/picture4.key   modified:   picture/picture4.pdf

  The  above limitation motivates us to propose a joint system that is capble of
  taking  advantagges  from  both  classification model and similarity model. We
  calle  it  SFC,  short  for  similairty model fused with classification Model,
  shown  in  Fig.  \ref{fig:ver.1.3}.  To  better train SFC, we further bring in
  multi-task   learning   \cite{caruana1993multitask,collobert2008unified}.  Our
  basic  idea  cames  from  two  stages. In the first stage, we use an auxiliary
  model  to  select top-K most possible labels for a input. This model can be an
  elasticsearch  \cite{divya2013elasticsearch}  or  a  text classification model
  trained on current task-specific chatbot data. In the second stage, we build a
  classification  model  whose  main modules are similarity models. Then, this
  structure  derives two goals to train towards, the classification loss for our
  application,  and the similarity loss for the main modules. When the two stages
  are  independently  optimized,  we call \emph{two-stage SFC}. We further find,
  the  the  quality  of  output  from  the  first  stage  might  limit the final
  performance  of  system,  since  the  candidate  labels for each user query is
  always  fixed  in  the  whole  multi-task  training  process. This observation
  motivates  us  to  further  improve  the  two-stage SFC into a joint training,
  called  \emph{joint  SFC}. In this way, the sentence pairs associated with the
  top-K  classes  also  become dynamic, and the text classification model in the
  first  stage  will also be further optimized to provide better top-K result by
  the final training loss.

@@

    \item To  better  train such a system, we bring in multi-task training to obtain
    expected performance.

    \item Experiment  results  from 4 public and 1 private short-text classification
    datasets,  show  that  our  proposed  SFC joint system can achieve significant
    improvements over strong baselines, especially in the low resource settings.
  \end{enumerate}

@@

  \section{Methodology}
  In  this  section,  we  will  describe our proposed system of similarity model fused with
  classification  model,  named  as  SFC, in detail. We come up with 2 different
  approaches  for  implementation  of SFC system, in which joint SFC is a
  further improvement of two-stage SFC for better performance.

  \subsection{two-stage SFC with multi-task learning}

@@

  \textbf{1.  positive  sentence pair sampling}: During the training process, we
  have  to first make sure that the ground truth class label $y_{i}$ for $x_{i}$
  is  within  the  candidate  class  label  set  $\mathcal{C'}$. If $y_{i}\notin
  \mathcal{C'}$,  we  will manually add $y_{i}$ into $\mathcal{C'}$ by replacing
  $c'_{K}$  with  $y_{i}$, since $c'_{K}$ is the least promising candidate label
  according  to  the  auxiliary  model  in stage 1. Afterwards, we will randomly
  sample  $P$  sentences  $\{x'_{0},x'_{1}, \dots, x'_{P}\}$ with the same class
  label as $y_{i}$ from dataset $\mathcal{D}$ to form the set of sentence pairs
  with   positive   label  $\mathcal{P'}_{i}=\{[(x_{i},  x'_{0}),  1],  [(x_{i},
  x'_{1}),   1],   \dots,   [(x_{i},   x'_{P}),  1]\}$,  where  $P$  is  also  a
  hyperparameter  set  by  us  that  controls  the number of sentences we should
  sample from each class.

  \textbf{2.  negative  sentence pair sampling}: As for negative sentence pairs,
  we  will  also  randomly  sample  $P$ sentences for each class in the negative
  candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The class labels
  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set of most confusing class
  labels comparing to the ground truth ${y_{i}}$, so we assume that the sentence
  pairs  with  negative  label grouped by user input query $x_{i}$ and sentences
  with class labels in $\mathcal{C'}\backslash \{y_{i}\}$ are strong adversarial
  negative  samples for sentence-pair similarity model that can help enhance the
  training  speed  and  performance.  According  to  the same method as positive
  sentence  pair sampling, we can obtain the set of sentence pairs with negative
  label  as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{0}), 0], [(x_{i}, x'_{1}), 0], ...,
  [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

@@

  \subsubsection*{
    stage 1: classification model for providing top-K candidate class labels
  }

@@

  \subsubsection*{
    stage 2: sentence-pair model for multi-task learning
  }

@@

  The 2 tasks for stage 2 training process are listed below.

  $\bullet$ \textbf{Task 1: Regular sentence-pair similarity task}

@@

  $\bullet$ \textbf{Task 2: Classification task on top-K classes based on sentence-pair model}

@@

  Therefore,  the  overall  loss  function for multi-task learning in stage 2 is
  shown in Equ. \ref{eq: two-stage SFC loss}. The training objective of stage 2 is
  to  minimize  the  weighted  sum  of task-specific losses. Here $\alpha_S$ and
  $\alpha_K$  are  weights  of  task  1  and  task  2  respectively,  which  are
  $\alpha_S=0.125$ and $\alpha_K=0.875$ in our setting. The reason for giving task 2 relatively more weights is that according to the ablation experiments we did in Section \ref{sec:exp}, task 2 contributes more to the overall performance.

  \begin{align}
    \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K
    \label{eq: two-stage SFC loss}
  \end{align}

  \begin{table*}
    \begin{centering}
      \scalebox{0.9}{
      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        \textbf{Dataset} & \textbf{Domain} & \textbf{\#Class} & \textbf{\#Samples} & \textbf{\#Samples/Class} & \textbf{Experimental Settings}\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
        \hline 
        HWU64 & intention detection & 64 & [320, 640, 960, 1280, 1920, 3200] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        CLINC150 & intention detection & 150 & [750, 1500, 2250, 3000, 4500, 7500] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        BANKING77 & intention detection & 77 & [385, 770, 1155, 1540, 2310, 3850] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
      \end{tabular}
      }
    \par
    \end{centering}
    \caption{Key statistics for experimental datasets and few shot settings}

@@

  \subsection{joint SFC with joint model structure}
  Despite  the  outstanding experimental result of two-stage SFC on various tasks,
  we still find a limitation brought by the two-stage system structure. In two-stage
  SFC, Stage 1 and stage 2 are separate, so they do not interact with each other
  during  training.  In  this  case,  the  performance  of stage 1 can limit the
  potential  of  stage  2, meanwhile, stage 2 also cannot give training feedback
  back to stage 1 for fine-tuning.

@@

  There  are  3  tasks in total during the training process of joint SFC,
  shown  in  Fig. \ref{fig:framework}, and the overall loss is also the weighted
  sum  of  all  the task-specific losses, shown in Equ. \ref{eq: hierachical SFC
  loss}. Here, $\alpha_S$, $\alpha_K$ and $\alpha_C$ are the weights for task 1:
  sentence-pair  similarity,  task  2:  top-K classification and task 3: general
  single   sentence  classification  respectively,  which  are  $\alpha_S=0.25$,
  $\alpha_K=0.25$ and $\alpha_C=0.5$ in our setting.

  \begin{align}
    \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K + \alpha_C \mathcal{L}^C
    \label{eq: hierachical SFC loss}
  \end{align}

@@

  $\bullet$ \textbf{Full-scale real-world task-specific faq chatbot dataset}

@@

  $\bullet$ \textbf{Subset of customer review dataset for extreme classification}

@@

  $\bullet$ \textbf{Subset of recently-released intention detection datasets}.

@@

  \subsection{Baselines}
  We choose 3 groups of representative baselines:

@@

>>>>

  The  above limitation motivates us to propose a joint system that is capble of
  taking  advantagges  from  both  classification model and similarity model. We
  calle  it  SFC,  short  for  similairty model fused with classification Model,
  shown  in  Fig.  \ref{fig:ver.1.3}.  To  better train SFC, we further bring in
  multi-task   learning   \cite{caruana1993multitask,collobert2008unified}.  Our
  basic  idea  cames  from  two  stages. In the first stage, we use an auxiliary
  model  to  select top-K most possible labels for a input. This model can be an
  elastic search  \cite{divya2013elasticsearch}  or  a  text classification model
  trained on current task-specific chatbot data. In the second stage, we build a
  classification  model  whose  main modules are similarity models. Then, this
  structure  derives two goals to train towards, the classification loss for our
  application,  and the similarity loss for the main modules. When the two stages
  are  independently  optimized,  we call \emph{two-stage SFC}. We further find,
  the  the  quality  of  output  from  the  first  stage  might  limit the final
  performance  of  system,  since  the  candidate  labels for each user query is
  always  fixed  in  the  whole  multi-task  training  process. This observation
  motivates  us  to  further  improve  the  two-stage SFC into a joint training,
  called  \emph{joint  SFC}. In this way, the sentence pairs associated with the
  top-K  classes  also  become dynamic, and the text classification model in the
  first  stage  will also be further optimized to provide better top-K result by
  the final training loss.

@@

    \item To  better  train such a system, we bring in multi-task training to obtain
    reasonable performance.

    \item Experiment  results  from 4 public and 1 private short-text classification
    datasets,  show  that  our  proposed  SFC joint system can achieve significant
    and consistent improvements over strong baselines, especially in the low resource settings.
  \end{enumerate}

@@

  \section{Methodology}
  In  this  section,  we  will  describe our proposed system of similarity model
  fused  with  classification  model,  named  as  SFC, in detail. We propose two
  approaches  to  implement  the  SFC idea, two-stageg SFC, and its more compact
  version called joint-SFC.

  \subsection{Two-stage SFC}

@@

  The  first step is \emph{positive sentence pair sampling}. During the training
  process,  we have to first make sure that the ground truth class label $y_{i}$
  for  $x_{i}$  is  within  the  candidate  class  label  set $\mathcal{C'}$. If
  $y_{i}\notin  \mathcal{C'}$,  we will manually add $y_{i}$ into $\mathcal{C'}$
  by  replacing  $c'_{K}$  with  $y_{i}$,  since $c'_{K}$ is the least promising
  candidate  label  according  to the auxiliary model in stage 1. Afterwards, we
  will randomly sample $P$ sentences $\{x'_{0},x'_{1}, \dots, x'_{P}\}$ with the
  same  class  label  as  $y_{i}$  from dataset $\mathcal{D}$ to form the set of
  sentence  pairs  with positive label $\mathcal{P'}_{i}=\{[(x_{i}, x'_{0}), 1],
  [(x_{i},  x'_{1}),  1],  \dots,  [(x_{i},  x'_{P}), 1]\}$, where $P$ is also a
  hyperparameter  set  by  us  that  controls  the number of sentences we should
  sample from each class.

  The  second  step  is  \emph{negative sentence pair sampling}: As for negative
  sentence  pairs,  we will also randomly sample $P$ sentences for each class in
  the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
  class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
  confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
  that  the  sentence  pairs  with  negative  label  grouped by user input query
  $x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
  are  strong  adversarial  negative  samples for sentence-pair similarity model
  that  can  help  enhance  the training speed and performance. According to the
  same  method  as  positive  sentence  pair  sampling, we can obtain the set of
  sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{0}),
  0], [(x_{i}, x'_{1}), 0], ..., [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

@@

  \subsubsection*{
    Stage 1: classification model for providing top-K candidate class labels
  }

@@

  \subsubsection*{Stage 2: sentence-pair model for multi-task learning}

@@

  \subsection{Multitask based training}
  The 2 tasks for stage 2 training process are listed below.

  \subsubsection*{Task 1: Regular sentence-pair similarity task}

@@

  \subsubsection*{Task   2:  Classification  task  on  top-K  classes  based  on
  sentence-pair model}

@@

  Therefore,  the  overall  loss  function for multi-task learning in stage 2 is
  shown in Equ. \ref{eq:two-stage_SFC_loss}. The training objective of stage 2 is
  to  minimize  the  weighted  sum  of task-specific losses. Here $\alpha_S$ and
  $\alpha_K$  are  weights  of  task  1  and  task  2  respectively.
  which  are
  $\alpha_S=0.125$ and $\alpha_K=0.875$ in our setting.

  \begin{align}
    \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K
    \label{eq:two-stage_SFC_loss}
  \end{align}

  \begin{table*}
    \begin{centering}
      \scalebox{0.9}{
      \begin{tabular}{|c|c|c|c|c|c|}
        \hline 
        \textbf{Dataset} & \textbf{Domain} & \textbf{\#Class} & \textbf{\#Samples} & \textbf{\#Samples/Class} & \textbf{Experimental Settings}\tabularnewline
        \hline 
        ITG & FAQ chatbot & 228 & 3938 & 17 & 3-fold\tabularnewline
        \hline 
        Amazon-670k & Product review & 250 & 1942 & 8 & class sampling\tabularnewline
        \hline 
        HWU64 & intention detection & 64 & [320, 640, 960, 1280, 1920, 3200] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        CLINC150 & intention detection & 150 & [750, 1500, 2250, 3000, 4500, 7500] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
        BANKING77 & intention detection & 77 & [385, 770, 1155, 1540, 2310, 3850] & [5, 10, 15, 20, 30, 50] & data sampling\tabularnewline
        \hline 
      \end{tabular}
      }
    \par
    \end{centering}
    \caption{Statistics for all datasets and few shot settings\footnote{All
    settings of four public datasets would be released to github soon}.}

@@

  \subsection{Joint SFC with joint model structure}
  In  two-stage  SFC,  Stage 1 and stage 2 are separate, so they do not interact
  with  each other during training. In this case, the performance of stage 1 can
  limit  the  potential of stage 2, meanwhile, stage 2 also cannot give training
  feedback back to stage 1 for fine-tuning.

@@

  There  are 3 tasks in total during the training process of joint SFC, shown in
  Fig. \ref{fig:framework}, and the overall loss is also the weighted sum of all
  the   task-specific  losses,  shown  in  Equ.  \ref{eq:joint_SFC_loss}.  Here,
  $\alpha_S$,   $\alpha_K$   and   $\alpha_C$   are  the  weights  for  task  1:
  sentence-pair  similarity,  task  2:  top-K classification and task 3: general
  single sentence classification respectively.
  \begin{align}
    \mathcal{L} = \alpha_S \mathcal{L}^S + \alpha_K \mathcal{L}^K + \alpha_C \mathcal{L}^C
    \label{eq:joint_SFC_loss}
  \end{align}

@@

  \textbf{Real-world task-specific faq chatbot dataset}

@@

  \textbf{Customer review dataset for extreme classification}

@@

  \textbf{Recently-released intention detection datasets}.

@@

  \subsection{Baselines}
  We choose 3 groups of typical baselines:

@@

  In  two-stage  SFC, in equ. \ref{eq:two-stage_SFC_loss} the  weights of two tasks are set as $\alpha_S=0.125$ and
  $\alpha_K=0.875$  in our setting. The reason for giving task 2 relatively more
  weights  is  that  according  to  the  ablation  experiments we did in Section
  \ref{sec:exp},  task  2  contributes more to the overall performance. In joint
  SFC,  in equ. \ref{eq:joint_SFC_loss}, the  weights  of three tasks are set as $\alpha_S=0.25$, $\alpha_K=0.25$
  and $\alpha_C=0.5$ in our setting.
